---
title: "Latent node activation analysis"
author: "Alexander Titus"
output:
  html_document: default
  html_notebook: default
---



# Setup the environment
```{r}
### Setup the environment we need ###
# install.packages('glmnet')
# install.packages('data.table')
require(glmnet)
require(data.table)

#dir = '/Users/alexandertitus/Documents/github/DNAm_data_generation/code'
dir = 'C:/Users/atitus/Documents/github/DNAm_data_generation/results'
setwd(dir)
```


# Load all the data and make variables
```{r}
### Load our data to work with ###
vae.file = '../results/encoded_methyl_onehidden_warmup_batchnorm_300K-100.tsv'
BRCA.covFile = '../BRCAtarget_covariates.csv'
tSNE.file = '../results/vae_tsne_out_300K-100_3d.tsv'

tSNE_features = data.frame(fread(tSNE.file), row.names=1)
colnames(tSNE_features) = tSNE_features[1, ]
tSNE_features = tSNE_features[2:nrow(tSNE_features), ]
tSNE_features$Basename = rownames(tSNE_features)

vae_features = data.frame(fread(vae.file), row.names=1)
colnames(vae_features) = vae_features[1, ]
vae_features = vae_features[2:nrow(vae_features), ]
vae_features$Basename = rownames(vae_features)

BRCA.covs = data.frame(fread(BRCA.covFile), row.names=1)

full.data = merge(vae_features, BRCA.covs, by='Basename')
full.data = merge(full.data, tSNE_features, by='Basename')
full.data$BasalVother = ifelse(full.data$PAM50.RNAseq == "Basal", 1, 0)
full.data$NormalVother = ifelse(full.data$PAM50.RNAseq == "Normal", 1, 0)
full.data$Her2Vother = ifelse(full.data$PAM50.RNAseq == "Her2", 1, 0)
full.data$LumAVother = ifelse(full.data$PAM50.RNAseq == "LumA", 1, 0)
full.data$LumBVother = ifelse(full.data$PAM50.RNAseq == "LumB", 1, 0)
full.data$LumVother = ifelse(full.data$PAM50.RNAseq == "LumA" | 
                                     full.data$PAM50.RNAseq == "LumB", 1, 0)
```


# Generate train and test sets
```{r}
### Modeling ###
#########################
# Split data
set.seed(1234)
train.prob = rbinom(nrow(full.data), 1, 0.5)
full.data$TranVtest = train.prob
```


# Normal classification

## 3 dimensional analysis
```{r}
#########################
# Build model - logistic regression "Classify as normal" - 3D
temp.data = full.data[, c('NormalVother', '1.y', '2.y', '3.y')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]

model <- glm(factor(NormalVother) ~., family=binomial(link='logit'), data=train.data) 
summary(model)

# Testing the model
fitted.results <- predict(model,newdata=test.data,type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$NormalVother)
normal3dAccuracy = 1-misClasificError
print(paste('3D Accuracy = ',normal3dAccuracy))
```


## 2 dimensional analysis
```{r}
#########################
# Build model - logistic regression "Classify as normal" - 2D
temp.data = full.data[, c('NormalVother', '1.y', '3.y')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]

model <- glm(factor(NormalVother) ~., family=binomial(link='logit'), data=train.data) 
summary(model)

#########################
# Testing the model
fitted.results <- predict(model,newdata=test.data,type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$NormalVother)
normal2dAccuracy = 1-misClasificError
print(paste('2D Accuracy = ',normal2dAccuracy))
```


# Luminal any classification
## 3D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 3D
temp.data = full.data[, c('LumVother', '1.y', '2.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumVother', '1.y', '2.y', '3.y')]
test.data = test.data[, c('LumVother', '1.y', '2.y', '3.y')]

model <- glm(factor(LumVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumVother)
print(paste('Accuracy',1-misClasificError))
```

## 2D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 2D
temp.data = full.data[, c('LumVother', '1.y', '2.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumVother', '1.y', '2.y')]
test.data = test.data[, c('LumVother', '1.y', '2.y')]

model <- glm(factor(LumVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumVother)
print(paste('Accuracy',1-misClasificError))
```


# Luminal A classification
## 3D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 3D
temp.data = full.data[, c('LumAVother', '1.y', '2.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumAVother', '1.y', '2.y', '3.y')]
test.data = test.data[, c('LumAVother', '1.y', '2.y', '3.y')]

model <- glm(factor(LumAVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumAVother)
print(paste('Accuracy',1-misClasificError))
```

## 2D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 2D
temp.data = full.data[, c('LumAVother', '1.y', '2.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumAVother', '1.y', '2.y')]
test.data = test.data[, c('LumAVother', '1.y', '2.y')]

model <- glm(factor(LumAVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumAVother)
print(paste('Accuracy',1-misClasificError))
```

# Luminal B classification
## 3D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 3D
temp.data = full.data[, c('LumBVother', '1.y', '2.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumBVother', '1.y', '2.y', '3.y')]
test.data = test.data[, c('LumBVother', '1.y', '2.y', '3.y')]

model <- glm(factor(LumBVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumBVother)
print(paste('Accuracy',1-misClasificError))
```

## 2D
```{r}
#########################
# Build model - logistic regression "Classify as Luminal" - 2D
temp.data = full.data[, c('LumBVother', '1.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('LumBVother', '1.y', '3.y')]
test.data = test.data[, c('LumBVother', '1.y', '3.y')]

model <- glm(factor(LumBVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$LumBVother)
print(paste('Accuracy',1-misClasificError))

# Basal-like classification

## 3D
```{r}
#########################
# Build model - logistic regression "Classify as Basal" - 3D
temp.data = full.data[, c('BasalVother', '1.y', '2.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('BasalVother', '1.y', '2.y', '3.y')]
test.data = test.data[, c('BasalVother', '1.y', '2.y', '3.y')]

model <- glm(factor(BasalVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$BasalVother)
basalAccuracy = 1-misClasificError
print(paste('Accuracy = ', basalAccuracy))
```

## 2D
```{r}
#########################
# Build model - logistic regression "Classify as Basal" - 2D
temp.data = full.data[, c('BasalVother', '1.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('BasalVother', '1.y', '3.y')]
test.data = test.data[, c('BasalVother', '1.y', '3.y')]

model <- glm(factor(BasalVother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$BasalVother)
basalAccuracy = 1-misClasificError
print(paste('Accuracy = ', basalAccuracy))
```

# Her2 classification

## 3D
```{r}
#########################
# Build model - logistic regression "Classify as Her2" - 3D
temp.data = full.data[, c('Her2Vother', '1.y', '2.y', '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('Her2Vother', '1.y', '2.y', '3.y')]
test.data = test.data[, c('Her2Vother', '1.y', '2.y', '3.y')]

model <- glm(factor(Her2Vother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$Her2Vother)
Her2Accuracy = 1-misClasificError
print(paste('Accuracy = ', Her2Accuracy))
```

## 2D
```{r}
#########################
# Build model - logistic regression "Classify as Her2" - 2D
temp.data = full.data[, c('Her2Vother',  '3.y', 'PAM50.RNAseq')]

train.data = temp.data[train.prob == 1, ]
test.data = temp.data[train.prob == 0, ]
table(train.data$PAM50.RNAseq);table(test.data$PAM50.RNAseq) 

train.data = train.data[, c('Her2Vother',  '3.y')]
test.data = test.data[, c('Her2Vother',  '3.y')]

model <- glm(factor(Her2Vother) ~., family=binomial(link='logit'), data=train.data) 

summary(model)

# Testing the model
fitted.results <- predict(model, newdata=test.data, type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)
misClasificError <- mean(fitted.results != test.data$Her2Vother)
Her2Accuracy = 1-misClasificError
print(paste('Accuracy = ', Her2Accuracy))
```


# Correlation testing
```{r}
#########################
# Latent activation correlation with age and survival
results = cor.test(BRCA.covs$age.Dx, vae_features$`1`)

w = sapply(c(1:100), function(x) cor(BRCA.covs[,1], vae_features[,x], method = "pearson", use = 'pairwise.complete.obs') )
w
results$estimate; results$p.value
```





